tl20-05_berghei_pir_DESEQ

### 2021-09-17 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Remove redundant comparisons and Change minReplicatesForReplace in DESEQ

```{r deseq-ing}
dds.deseq <- DESeq(dds)
resultsNames(dds.deseq)
```

@@

```{r gamMale_v_asexMix}
coef <- "stage_Gam.Male_vs_Asex.Mixed"
dds.deseq.gamMale_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  type = "apeglm")

dds.deseq.gamMale_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% pir_info$Geneid,] %>% 
  as.data.frame() %>% 
  mutate(subfam = pir_info$clade[match(rownames(.), pir_info$Geneid)]) %>% 
  arrange(padj)

gamMale_v_asexMixed_df <- as.data.frame(
  dds.deseq.gamMale_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0317181'),]
) %>% 
  mutate(contrast = coef) %>% 
  rownames_to_column(var = 'Geneid')
```

```{r}
coef <- "stage_Liv.Meroz_vs_Asex.Mixed"
res <- results(dds.deseq,
               name = 'stage_Liv.Meroz_vs_Asex.Mixed',
               cooksCutoff = FALSE)
dds.deseq.livMeroz_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  res = res,
                                                  type = "apeglm")

livMeroz_v_asexMixed_df <- as.data.frame(
  dds.deseq.livMeroz_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0100500',
                                                                                              'PBANKA_0600031'),]) %>%  
  mutate(contrast = coef) %>% 
  rownames_to_column(var = 'Geneid')
```

```{r asexRing_v_asexMixed}
coef <- 'stage_Asex.Ring_vs_Asex.Mixed'

dds.deseq.asexRing_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  type = "apeglm")
asexRing_v_asexMixed_df <- as.data.frame(
  dds.deseq.asexRing_v_asexMixed.lfcShr[rownames(dds.deseq.asexRing_v_asexMixed.lfcShr) %in% c('PBANKA_1300200'),]
) %>% 
  mutate(contrast = coef) %>% 
    rownames_to_column(var = 'Geneid')
```

```{r}
dds$stage <- relevel(dds$stage, ref = 'Gam-Male')
dds.maleGam.deseq <- DESeq(dds)
resultsNames(dds.maleGam.deseq)
```

```{r}
coef <- "stage_Liv.Meroz_vs_Gam.Male"
res <- results(dds.maleGam.deseq,
               name = 'stage_Liv.Meroz_vs_Gam.Male',
               cooksCutoff = FALSE)
dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr <- lfcShrink(dds.maleGam.deseq, 
                                                         coef = coef,
                                                         res= res,
                                                         type = "apeglm")

(livMeroz_v_gamMale_df <- as.data.frame(
  dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr[rownames(dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr) %in% c('PBANKA_0317181',
                                                                                                           'PBANKA_0100500',
                                                                                                           'PBANKA_0600031'),]
  ) %>%  
  mutate(contrast = coef) %>% 
    rownames_to_column(var = 'Geneid'))
```

```{r}
(deseq_stats <- bind_rows(livMeroz_v_gamMale_df, 
          as.data.frame(gamMale_v_asexMixed_df),
          as.data.frame(livMeroz_v_asexMixed_df)) %>% 
  mutate(FDR = p.adjust(pvalue, 
                        method = 'fdr')) %>% 
  select(-c(lfcSE, 
            padj, 
            pvalue)))
```

@@

shrunk.list <- lapply(comparisons, 
                      function(comparison){
                        res <- results(
                          dds.deseq,
                          contrast = c("stage",comparison[1],comparison[2])
                        )
                        shrunk <- lfcShrink(dds.deseq, 
                                            res = res,
                                            type = "ashr") %>% 
                          as.data.frame %>% 
                          rownames_to_column(var = 'Geneid') %>% 
                          mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
                          filter(Geneid %in% pir_id,
                                 !is.na(padj))
                        return(shrunk)
                      })
names(shrunk.list) <- unlist(lapply(comparisons, 
                                    function(comparison) paste0(comparison[1],'_',comparison[2])))
sig_pirs <- bind_rows(shrunk.list) %>% 
  mutate(fdr = p.adjust(pvalue)) %>% 
  arrange(subfam, Geneid) %>% 
  filter(fdr < 0.05)
sig_pirs_order <- unique(sig_pirs$Geneid)
sig_pirs_cast <- dcast(sig_pirs,
                  comp ~ Geneid,
                  value.var = 'log2FoldChange') %>% 
  column_to_rownames(var = 'comp')
sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
                              r_order_vector),
                     sig_pirs_order]

>>>>

```{r deseq-ing}
dds.deseq <- DESeq(dds,
                   minReplicatesForReplace = 3)
resultsNames(dds.deseq)
```

@@

@@

shrunk.list <- lapply(comparisons, 
                      function(comparison){
                        res <- results(
                          dds.deseq,
                          contrast = c("stage",comparison[1],comparison[2])
                        )
                        shrunk <- lfcShrink(dds.deseq, 
                                            res = res,
                                            type = "ashr") %>% 
                          as.data.frame %>% 
                          rownames_to_column(var = 'Geneid') %>% 
                          mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
                          filter(Geneid %in% pir_id,
                                 !is.na(padj))
                        return(shrunk)
                      })
names(shrunk.list) <- unlist(lapply(comparisons, 
                                    function(comparison) paste0(comparison[1],'_',comparison[2])))
sig_pirs <- bind_rows(shrunk.list) %>% 
  mutate(fdr = p.adjust(pvalue),
         subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
  arrange(subfam, Geneid) %>% 
  filter(fdr < 0.05)
sig_pirs_order <- unique(sig_pirs$Geneid)
sig_pirs_cast <- dcast(sig_pirs,
                  comp ~ Geneid,
                  value.var = 'log2FoldChange') %>% 
  column_to_rownames(var = 'comp')
sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
                              r_order_vector),
                     sig_pirs_order]

sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031')
